### 2021-11-03 paper/report/summary_stats.Rmd
initial commit of summary stats R files

---
title: "Second Year Paper Summary Statistics"
author: "Aaron Watt"
email: aaron@acwatt.net
date: "2021-11-08"
output:
  pdf_document:
    toc: true
    toc_depth: 2
header-includes:
   - \usepackage{dcolumn}
   - \usepackage{amsmath}
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'ARE_213_PS2a.pdf')) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<!--
R version 3.6.3 (2020-02-29)
Purpose of script: Second Year Paper summary statistics

Notes: Need to ps2a.R from ps2a folder (without Rstudio being started first)
       if Rstudio is already started, the working directory will not be set
       to the ps2a/ folder

       \usepackage{dcolumn}: dcolumn is needed in latex compilation 
-->

```{r Settings, echo=FALSE}
# stargazer table type (html, latex, or text)
# Change to latex when outputting to PDF, html when outputting to html
table_type = "text"
```

```{r packages, results='hide', message=FALSE, echo=T}
library(tidyverse)
library(foreign)
library(stargazer)
```