### 2021-12-11 paper/5min-presentation/feedback/shreya_model.Rmd
addign shreya s feedback

---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}

```

Assume $X_{m,y}$ is AR(1) + random normal distribution
```{r}
library(tidyverse)
library(magrittr)
set.seed(1)
# Define DGP parameters
months = 12
years = 2
households = 4
starting_x = 20
rho = 0.8
beta0 = 10
beta1 = 1  # for every dollar of subsidy, we want one more unit of gas consumed
# Generate household, month, year shocks (abosorbed by gamma fixed effects)
e_h = comprehenr::to_vec(for(h in 1:(households)) rnorm(1))
e_m = comprehenr::to_vec(for(h in 1:(months)) rnorm(1, mean=0, sd=2))
e_y = comprehenr::to_vec(for(h in 1:(years)) rnorm(1, mean=-2, sd=2))
# Generate the AR(1) subsidies
X = arima.sim(model = list(order = c(1, 0, 0), ar = rho), n = months*years, sd=6) + 20
plot(X)
# Create dataframe
df = data.frame(h = factor(comprehenr::to_vec(for (h in 1:households) for (y in 1:years) for (m in 1:months) h)),
                y = factor(comprehenr::to_vec(for (h in 1:households) for (y in 1:years) for (m in 1:months) y)),
                m = factor(comprehenr::to_vec(for (h in 1:households) for (y in 1:years) for (m in 1:months) m)),
                X = comprehenr::to_vec(for (h in 1:households) X),
                e_h = rep(e_h, each = months*years),
                e_y = rep(e_y, each = months, times=households),
                e_m = rep(e_m, times = households*years),
                e_hmy = rnorm(n = households*months*years, mean = 0, sd = 0.5)) %>%
    mutate(ym = paste0(y, '-', format(m, width=2)),
           # Generate Outcomes Y
           Y = beta0 + beta1*X + e_h + e_m + e_y + e_hmy)

df %>%
    ggplot() + geom_point(aes(x=X, y=Y, color = ym))

# Regression
df %>%
    lm(Y ~ X + h + m + y, data=.)
e_h
e_m
e_y
```