### 2021-12-15 paper/5min-presentation/5min-presentation.Rmd
Made some chunks all appear at the same time.

---
title: "Estimation and Implications of Bias in EPA Pollution Measurement"
author: Aaron C Watt
date: December 2, 2021
output:
  beamer_presentation:
    includes:
      in_header: preamble.tex
classoption: "aspectratio=169"
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'acWatt_purpleair_5min-pres_2021-12-15.pdf')) })
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
# Had to run tinytex::install_tinytex() to use XeLaTeX if using YGbeamer.sty
```

```{r packages, include=F}
library(tidyverse)
library(docstring)
library(ggplot2)
# library(mapdata)
# Requires rgdal package to be installed, requires gdal to be installed on computer
# Ubuntu 20.04: sudo aptitude install gdal-bin proj-bin libgdal-dev libproj-dev
```

# Motivation
## Clean Air

\begin{itemize}
\item<2-> The Clean Air Act (1970) established National Ambient Air Quality Standards
(NAAQS) for US counties
\item<2-> Either "attainment" or "non-attainment", \color{red}large penalties and costly restrictions
\item<2-> Minimum requirement of 75\% of readings, per quarter
\item<2-> Air quality can change quickly
\item<2-> Monitor shutoffs are common
\end{itemize}

<!-- - Annual and Daily standards that both must be met. -->

<!-- - Regulatory penalties levied on non-attainment areas can be large. -->

## Research Questions
\begin{itemize}
\item<3-> How biased are EPA monitor-based measures of local air quality?
\item<3-> Does this bias significantly change NAAQS attainment status?
\end{itemize}
<!-- - How biased are self-reported EPA pollution estimates? -->
<!-- - How biased is *missing* air pollution data from self-reporting US EPA monitors? -->

# Context
<!-- :
     latex_engine: xelatex
header-includes:
   - \usetheme[style=dark]{Nord}
   - \usepackage{YGbeamer}
(i) an interesting, well-posed question; 
(ii) an adequate model, such that the question makes sense within the context of
the model; 
(iii) an identification strategy, such that (again within the context of the 
model) the question can be answered; 
“What is your identification strategy?” = “What research design (and assumptions) are you using?”
(iv) a set of proposed hypothesis tests or statistics which map your
data into falsifiable statements that can provide answers to your question. You
should /not/ provide any results; this presentation is about empirical strategy,
not about data-dependent outcomes.
-->

## Previous Works in EPA Pollution Monitors

\uncover<2->{**Satellite Data:**}
\begin{itemize}
\setlength{\itemsep}{0em}
\item<2-> Grainger et al. 2017
\item<2-> Sullivan, Krupnick 2018
\item<2-> Fowlie, Rubin, Walker 2019
\item<2-> Zou 2021
\end{itemize}

\uncover<3->{**Location Pollution Alerts:**
\begin{itemize}
\setlength{\itemsep}{0em}
\item<3-> Mu, Rubin, Zou 2021
\end{itemize}

<!-- - Spotting Strategic Behavior in EPA Monitor Shutoffs: Mu, Rubin, Zou 2021 -->
<!-- [ put more here on strategic behavior evidence that motivates the interest in measuring -->
<!-- the size / distribution of bias in missing data/] -->

<!-- - Strategic decisions of monitor locations: Grainger et al. 2017 -->

<!-- - Using satellite data to fill gaps in air pollution monitoring: Sullivan, Krupnick 2018 -->

<!-- - Prediction error is significant in satellite-based estimates of air pollution: Fowlie, Rubin, Walker 2019 -->

\uncover<4->{## This project}
\begin{itemize}
%\setlength{\itemsep}{0em}
\item<4-> Using new consumer-based pollution monitors (PurpleAir).
\item<4-> Avoids using satellite estimates (has been shown to have significant error).
\item<4-> Focus on EPA pollution data that is missing *in time*, specifically PM2.5
\end{itemize}

<!-- - Understanding the distribution of bias in missing California air pollution data using consumer-based pollution -->
<!-- monitors that have low error near their location. -->

<!-- [why California? Maybe should inlcude all EPA monitor - months that have at least one PA monitor within 10 miles] -->

<!-- [insert maps of California EPA and PA monitors, timelaps GIF? Timeline of adoption] -->

<!-- [insert pictures of PA outdoor monitors] -->

```{r geography intersection functions, include=F}
library(sf)
library(spData)

lonlat_to_state <- function(df, lon_col, lat_col,
                            states = spData::us_states,
                            name_col = "NAME") {
  #' State names for lon, lat vectors
  #' 
  #' @description Return a vector of state names for the lon-lat vectors
  #' 
  #' @param df: a data.frame containing lon lat pairs
  #' 
  #' @param lon_col: A character of the column name containing longitude
  #' 
  #' @param lat_col: A character of the column name containing latitude
  #'
  #' @param states:   An sf MULTIPOLYGON object with 50 states plus DC.
  #'
  #' @param name_col: Name of a column in `states` that supplies the states'
  #'           names.
  #' https://stackoverflow.com/a/8751965/16660792
  ## Convert points data.frame to an sf POINTS object
  pointsDF = data.frame(x = df[, lon_col], y = df[, lat_col])
  pts <- st_as_sf(pointsDF, coords = 1:2, crs = 4326)

  ## Transform spatial data to some planar coordinate system
  ## (e.g. Web Mercator) as required for geometric operations
  states <- st_transform(states, crs = 3857)
  pts <- st_transform(pts, crs = 3857)

  ## Find names of state (if any) intersected by each point
  state_names <- states[[name_col]]
  ii <- as.integer(st_intersects(pts, states))
  state_names[ii]
}

lonlat_country_filter <- function(df, lon_col, lat_col,
                                  country='United States',
                                  name_col='name_long') {
  #' Filter for lat-lons in a specific country
  #' 
  #' @description Return the dataframe filtered to inlcude only US points
  #' 
  #' @param df: a data.frame containing lon lat pairs
  #' 
  #' @param lon_col: A character of the column name containing longitude
  #' 
  #' @param lat_col: A character of the column name containing latitude
  #'
  #' @param country: A character of the country, matching name_col from spData::world
  #'
  #' @param name_col: A character of the column name containing the country name,
  #' from spData::world
  #'
  ## Restrict to county boundaries
  country = spData::world %>%
    filter(!!as.symbol(name_col) == country)
  ## Convert points data.frame to an sf POINTS object
  pointsDF = data.frame(x = df[, lon_col], y = df[, lat_col])
  pts <- st_as_sf(pointsDF, coords = 1:2, crs = 4326)

  ## Transform spatial data to some planar coordinate system
  ## (e.g. Web Mercator) as required for geometric operations
  country <- st_transform(country, crs = 3857)
  pts <- st_transform(pts, crs = 3857)

  ## Find names of state (if any) intersected by each point
  ii <- as.integer(st_intersects(pts, country))
  df %>%
    mutate(include = ii) %>%
    filter(include == 1)
}

```

# Use PM2.5 Air Pollution Monitors to Predict Missing EPA Data

```{r get EPA sensor locations 2016, include=F, cache=T}
sensor_types = c(88101, 88501, 88502)
years = c()

# The below data files are only on Aaron's desktop. Need to create another data submodule to host data
# or host on s3 bucket and download from there
load("../../data/epa/epa_monitors/epa_monitors/data_from_api/88101/api_request_daily_88101_2016.Rdata")
df = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88101, NAAQS_used='yes')
load("../../data/epa/epa_monitors/epa_monitors/data_from_api/88501/api_request_daily_88501_2016.Rdata")
df2 = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88501, NAAQS_used='no')
load("../../data/epa/epa_monitors/epa_monitors/data_from_api/88502/api_request_daily_88502_2016.Rdata")
df3 = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88502, NAAQS_used='no')
epa_data = bind_rows(df, df2, df3) %>%
  filter(type == 88101) %>% 
  distinct(latitude, longitude, .keep_all = TRUE) %>%
  select(longitude, latitude, everything())

```

```{r get PA sensor locations all, include=F}
result = jsonlite::fromJSON("../../data/test_data/manual_webpage_download_USmap_2021-11-29.txt")
library(zoo)
# Get start and last_seen dates from all sensors near the US
pa_data2 = as.data.frame(result$data) %>%
  purrr::set_names(result$fields) %>%
  mutate(date_start = lubridate::as_datetime(date_created),
         date_last = lubridate::as_datetime(last_seen),
         month_start = as.yearmon(date_start),
         month_last = as.yearmon(date_last)) %>%
  select(longitude, latitude, everything()) %>%
  arrange(date_start) %>%
  lonlat_country_filter(lon_col='longitude', lat_col='latitude') %>%
  filter(private == 0,  # public sensor
         location_type == 0,  # outdoor sensor
         channel_state != 0)  # has at least one PM sensor
```

```{r Plot the locations, eval=F, include=F}
min_lat = min(pa_data$lat)
max_lat = max(pa_data$lat)
min_lon = min(pa_data$lon)
max_lon = max(pa_data$lon)
print(paste(max_lon, max_lat, min_lon, min_lat))
usa <- map_data('usa', projection='conic', parameters=c(min_lat))

b = 204/255
plot_usmap(fill='white', color='grey') +
  geom_point(data = usmap::usmap_transform(pa_data),   # Purple Air
             aes(x = lon.1, y = lat.1),
             color = rgb(0, b, b), alpha = 1, size=0.4) +
  geom_point(data = usmap::usmap_transform(epa_data),   # EPA
             aes(x = longitude.1, y = latitude.1),
             color = "red", alpha = 0.25, size=2) +
  labs(title = "US EPA & PurpleAir Pollution Monitors",
       subtitle = "Source: EPA, PurpleAir.comJan 1 to Jun 30 2016") +
  theme(legend.position = "right")

```

```{r Monitor Map with AK and HI, include=F}
# devtools::install_github("hrbrmstr/albersusa")
# sudo apt-get install libudunits2-dev
# install.packages('units')
# install.packages('sf')

# Composite map with AK & HI
usa_map = albersusa::usa_composite(proj="laea") %>%
  fortify(.)  # make it usable in ggplot2
epa_trans <- usmap::usmap_transform(epa_data)
pa_trans <- usmap::usmap_transform(pa_data2)

# Plot PA and EPA sensor locations on US map
ggplot() +
  geom_map(data=usa_map, map=usa_map,
                    aes(map_id=id),
                    color="#2b2b2b", size=0.1, fill=NA) +
  geom_point(data = pa_trans, 
             aes(x = longitude.1, y = latitude.1),  # Purple Air
             color = rgb(0, 204/255, 204/255), alpha = 1, size=0.4) +
  geom_point(data = epa_trans, aes(x = longitude.1, y = latitude.1),  # EPA
             color = "red", alpha = 0.25, size=2) +
  labs(title = "US EPA & PurpleAir Pollution Monitors",
       subtitle = "Source: EPA 2016, PurpleAir.com 2015-2021") +
  ggthemes::theme_map()

ggsave('monitor_map_2015-2021.png', width = 6, height = 4)

```

\vspace{-2em}

```{r Map and images, echo=F, warning=F, message=F, fig.width=12, fig.cap='test'}
p_pa = "../../docs/purpleair/PA_outdoor_monitor_border.png"
p_epa = "../../docs/epa/air-monitoring-site_border.jpg"
p_map = 'monitor_map_2015-2021.png'
library(png)
library(jpeg)
library(grid)
library(gridExtra)
img_map = rasterGrob(as.raster(readPNG(p_map)), interpolate = FALSE)
img1 <-  rasterGrob(as.raster(readPNG(p_pa)), interpolate = FALSE)
img2 <-  rasterGrob(as.raster(readJPEG(p_epa)), interpolate = FALSE)

lay = rbind(c(1,2),
            c(1,3))
grid.arrange(img_map, img1, img2,
             layout_matrix = lay,
             widths = c(3,1))
```

```{r animate purpleair adoption, eval=F, include=F}
##Source: https://github.com/thomasp85/gganimate
library(ggplot2)
library(gganimate)
library(gapminder)
## Warning: package 'gapminder' was built under R version 3.4.4
library(gifski)
## Warning: package 'gifski' was built under R version 3.4.4

ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) +
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) +
  ease_aes('linear')

month_counts = pa_data2 %>%
  group_by()

dates <- as.Date(df1$date, '%d/%m/%Y')
yr <- year(dates)
monyr <- as.yearmon(dates)
lst <- lapply(list(dates, yr, monyr), function(x) 
       transform(df1, Count=ave(seq_along(x), x, FUN= length)))
names(lst) <- paste0('newdf', seq_along(lst))
list2env(lst, envir=.GlobalEnv)

```

# Models

\begin{itemize}%[itemsep=1.5em]
\item<2> \textbf{Estimate missing pollution observations:} Use PurpleAir data to predict EPA pollution at missing times
\item<2> \textbf{Estimate bias of reported EPA pollution:} difference between predicted pollution at missing times and reported pollution at nonmissing times.
\item<2> \textbf{Estimate counties' counterfactual attainment status:} Include estimated missing pollution data.
\end{itemize}

# Models: Predictive model of each EPA monitor PM2.5 pollution

\[
\alert<2>{EPA_{i,t}} = \gamma_{i,0} + \sum\limits_{j\in J_i}\sum\limits_{k=1}^7\gamma_{j,k} 
\alert<3>{PA_{j,t}} \cdot \alert<4>{Winddir_{i,t,k}} + u_{i,t}
\]

\begin{itemize}
\item<1-> Analysis done at the month and quarter level; suppressing that subscript.
\item<1-> $t$ is a unique hour within a given month or quarter.
\item<1-> EPA monitor $i$ at time $t$ reads PM2.5 pollution \alert<2>{$EPA_{i,t}$}.
\item<1-> For each EPA monitor $i$, there are $J_i$ Purple Air monitors within a 10-mile radius.
\item<1-> Purple Air monitor $j\in J_i$ at time $t$ reads PM2.5 pollution \alert<3>{$PA_{j,t}$}.
\item<1-> $\alert<4>{Winddir_{i,t,k}} $ is a wind direction indicator; 1 if the prevailing wind near station $i$ at time $t$ is in the $k^{th}$ bucket (of 8 buckets).
\end{itemize}

<!-- [insert diagram of two EPA monitors with PA monitors surrounding them] -->

<!-- [insert diagram of one EPA monitor and surrounding PA monitors, with wind directions] -->

# Models: Hour-by-Day-of-week Bias of Missing EPA Monitor Pollution Data
\def\mm{\mathcal{M}}
\def\nn{\mathcal{N}}
\vspace{-1em}
\begin{align*}
\intertext{\textbf{Missingness Bias:}}
Bias_{i,h,d} &= \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}EPA_{i,t} 
- \frac{1}{|\mm_{i,h,d}|}\sum\limits_{t\in \mm_{i,h,d}}\widehat{EPA}_{i,t}\\
\text{where } \mm_{i,h,d}&= \{t: t\text{ is at hour $h$ and day $d$ and $EPA_{i,t}$ is Missing}\};\\
  \nn_{i,h,d}&= \{t: t\text{ is at hour $h$ and day $d$ and $EPA_{i,t}$ is Non-missing}\} \\
\uncover<2->{\intertext{We can also define the \textbf{algorithm bias}     as the Hour-by-Day-of-week prediction error}
\widetilde{Bias}_{i,h,d} &= \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}EPA_{i,t} 
- \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}\widehat{EPA}_{i,t} \\}
\uncover<3->{\intertext{We can also define $Bias_{j,h,d}$ and $\widetilde{Bias}_{j,h,d}$ for PA monitor $j$ (we'll come back to this).}}
\end{align*}

<!-- \begin{itemize}[<+->] -->
<!-- \item \begin{align*} -->
<!-- Bias_{i,h,d} &= \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}EPA_{i,t}  -->
<!-- - \frac{1}{|\mm_{h,d}|}\sum\limits_{t\in \mm_{h,d}}\widehat{EPA}_{i,t}\\ -->
<!-- \text{where } \mm_{i,h,d}&= \{t: t\text{ is at hour $h$ and day $d$ and $EPA_{i,t}$ is Missing}\};\\ -->
<!--   \nn_{i,h,d}&= \{t: t\text{ is at hour $h$ and day $d$ and $EPA_{i,t}$ is Non-missing}\} -->
<!-- \end{align*} -->

<!-- \item We can also define the algorithm bias as the Hour-by-Day-of-week prediction error -->
<!-- \[ -->
<!-- \widetilde{Bias}_{i,h,d} = \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}EPA_{i,t}  -->
<!-- - \frac{1}{|\mm_{h,d}|}\sum\limits_{t\in \mm_{h,d}}\widehat{EPA}_{i,t} -->
<!-- \] -->

<!-- \item We can also define $Bias_{j,h,d}$ and $\widetilde{Bias}_{j,h,d}$ for PA monitor $j$ (we'll come back to this). -->
<!-- \end{itemize} -->

# Models: County Attainment Status

\begin{align*}
Attain^{annual}_c &= 1 \text{ if \textbf{reported} annual average PM2.5 below 15.0 $\mu$g/m$^3$}^*\\[1em]
Attain^{daily}_c &= 1 \text{ if 98$^{th}$ percentile of \textbf{reported} daily average PM2.5 below 35 $\mu$g/m$^3$}^*\\[1em]
\uncover<2->{\widehat{Attain}^{annual}_c &= 1 \text{ if \textbf{predicted} annual average PM2.5 below 15.0 $\mu$g/m$^3$}^*}\\[1em]
\uncover<2->{\widehat{Attain}^{daily}_c &= 1 \text{ if 98$^{th}$ percentile of \textbf{predicted} daily average PM2.5 below 35 $\mu$g/m$^3$}^*}
\end{align*}
$^*$averaged over 3 years in NAAQS standard.
[fill in equations and thresholds]

# Identification

\uncover<2->{\textbf{Is the missing pollution data identified?}}

\begin{itemize}
\item<3-> Assumption: nearby PurpleAir monitors that are good predictors for EPA monitors
during non-missing times will also be good predictors during missing times.
\item<3-> Specifically, reasons for EPA data missingness are not correlated with missingness
  or measurement error in PurpleAir data
\end{itemize}

\uncover<4->{\textbf{Are attainment status changes identified?}}

\begin{itemize}
\item<5-> Yes: simple rule based on pollution concentrations.
\end{itemize}

# Proposed Statistical Test

- The $J_i$ group of PurpleAir sensors is (in a sense) a synthetic control for the EPA sensor $i$.

- **The question of bias in EPA monitor data can be stated**: is pollution concentration at
the monitor during reported times
significantly different from the concentration at
the monitor at missing times? Is the difference greater than random variations?

- Implies an permutation inference test for each EPA monitor $i$.

# Proposed Statistical Test (Is the bias larger than by random chance?)

```{r Create bar plots, eval=F, include=F}
library(RColorBrewer)

# create dummy data
b = rgb(0.7,0.7,0.7)
r = rgb(0.8,0.1,0.1)
mm = "Missing"
nn = "Reported"
data1 <- data.frame(
  hour=1:24,
  value=sapply(1:24, function (x) rnorm(1, 20-0.05*(x-14)^2, 2)),
  color=c(rep(b,15), rep(r,5), rep(b, 24-15-5)),
  missing=c(rep(nn,15), rep(mm,5), rep(nn, 24-15-5))
)

data2 <- data.frame(
  hour=1:24,
  value=sapply(1:24, function (x) rnorm(1, 20-0.05*(x-14)^2, 2)),
  color=c(rep(b,15), rep(r,5), rep(b, 24-15-5)),
  missing=sample(c())  c(rep(nn,15), rep(mm,5), rep(nn, 24-15-5))
)

# Uniform color
ggplot(data) +
  geom_col(aes(x=hour, y=value, fill=missing)) +
  scale_fill_manual(values = c(r,b)) +
  xlab("Hour of day") +
  ylab("PM2.5 Pollution") +
  ggtitle("EPA Monitor") +
  theme(text = element_text(size=20))
barplot(height=data$value, names=data$hour, col=data$color, cex.names = 1.4,
        cex.axis=1.5, cex=1.5, xlab="hour", ylab="PM2.5 Pollution", main="EPA Monitor")

```

- Calculate missingness bias and algorithm bias for the $i^{th}$ EPA monitor using $J_i$ PA monitors with $n_i$ missing hour observations.

- Pick PA monitor $j\in J_i$, temporarily remove random $n_i$ hour observations.

- Construct placebo synthetic control for PA monitor $j$ and predict $\widehat{PA}_{j,t}$.

- Calculate missingness bias and algorithm bias for PA monitor $j$: $Bias_{j,h,d}$

- Repeat for all PA sensors.

# Proposed Statistical Test (Is the bias larger than by random chance?)

- **Graphical test**: For EPA sensor $i$, compare graph of $Bias_{i,h,d}$ to placebo $Bias_{j,h,d}$ for $j\in J_i$.

- **Permutation inference p-value**:

  - Calculate sum of squared missingness bias and sum of squared algorithm bias for EPA sensor $i$ and PA sensors $j\in J_i$.
  \vspace{1em}
  - Ratio$_k$ = sum of squared missingness bias / sum of squared algorithm bias
  \vspace{1em}
  - p-value = $\frac{\text{\# of PA sensors in $i$'s radius with Ratio$_j$ larger than Ratio$_i$}}{\text{\# of PA sensors in $i$'s radius}}$

<!-- But how does this account for poor prediction of the monitor?
Because the denominator is larger if poor prediction in non-missing observations -->

# Extensions

- Use PurpleAir data to create population-weighted pollution measure $\implies$ counterfactual attainment.

- Welfare analysis based on attainment status changes and required reductions in pollution.

- Comparing county population-weighted PM2.5 pollution to EPA sensors to estimate location-based bias.

# Appendix A: PurpleAir monitor correction factor

![PurpleAir correction equation for EPA monitor PM2.5 (RH = relative humidity, also measured by PA monitor)](../../docs/purpleair/PA_correction_equation.png){width=50%}

Source: \url{https://www.epa.gov/air-sensor-toolbox/technical-approaches-sensor-data-airnow-fire-and-smoke-map}

\beamerdefaultoverlayspecification{}
# Appendix B: Data Plan
### Datasets

- Hourly PM2.5 Pollution data from California EPA pollution monitors (2015-2020)

- 2-minute PM2.5 Pollution data from California PurpleAir sensors, hourly averages taken

### Deadlines
- Dec. 5: PurpleAir is downloading/averaging on 4 AWS tiny linux instances, sending CSVs to S3 bucket
- Dec. 12: Proof of concept for 2 EPA sensors (Fresno, and [need to pick another low on Mu's list])
- Dec. 19: Data warehouse setup and transfer of existing Purple Air data

# Appendix B: Data Plan

### Data Wearhouse
- AWS Linux Cassandra database (noSQL columnar, designed for large queries of columns)
- Python pushes and pulls

### Storage costs
- ~ 30,000 sensors, 50 variables, 2 minute intervals, 5 years of data = 107.55 Terabytes
- Depending on method of storage: \$2,900 - \$13,300 per month
- Only storing hourly means and SD: \$4 - \$15 per month

# Appendix C: PurpleAir Takeup
![Valid Purple Air Monitor Locations, Contiguous United States](../../output/summary_stats/graphs/purple_air_sensor_cum_time_california.png){width=60%}

<!-- Extra
# Example
### Two EPA monitors:
- Fresno (Mu et al. identified as "interesting")
- other county low on the interesting list

# Data Plan
### Datasets

### Calendar

### Data Wearhouse
- AWS Linux Cassandra database (noSQL columnar, designed for large queries of columns)

### storage costs
- ~ 30,000 sensors, 50 variables, 2 minute intervals, 5 years of data = 107.55 Terabytes
- Depending on method of storage: \$2,900 - \$13,300 per month
- Only storing hourly means and std. deviations: \$4 - \$15 per month

-->