### 2022-03-06 pics/appendix/tables/attainment_counties.tex

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage{multirow}
% \usepackage{longtable}
% Note: It may be necessary to compile the document several times to get a multi-page table to line up properly
\begin{longtable}{@{}lllll@{}}
\caption{Current NAAQS Nonattainment Counties}
\label{tab:naaqs_attain_counties}\\
\toprule
\textbf{State} & \textbf{Area Name} & \textbf{\begin{tabular}[c]{@{}l@{}}EPA Designated \\ Nonattainment Counties\end{tabular}} & \textbf{FIPS} & \textbf{\begin{tabular}[c]{@{}l@{}}\# of monitors\\ in sample\end{tabular}} \\* \midrule
\endfirsthead
%
\endhead
%
\multicolumn{1}{|l|}{\multirow{14}{*}{CA}} & \multicolumn{1}{l|}{Imperial County, CA} & \multicolumn{1}{l|}{Imperial, CA (p)} & \multicolumn{1}{l|}{025} & \multicolumn{1}{l|}{} \\* \cmidrule(l){2-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{\multirow{8}{*}{\begin{tabular}[c]{@{}l@{}}San Joaquin Valley \\ Air Basin, CA\end{tabular}}} & \multicolumn{1}{l|}{Fresno, CA} & \multicolumn{1}{l|}{019} & \multicolumn{1}{l|}{2} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Kern, CA (p)} & \multicolumn{1}{l|}{029} & \multicolumn{1}{l|}{1} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Kings, CA} & \multicolumn{1}{l|}{031} & \multicolumn{1}{l|}{2} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Madera, CA} & \multicolumn{1}{l|}{039} & \multicolumn{1}{l|}{1} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Merced, CA} & \multicolumn{1}{l|}{047} & \multicolumn{1}{l|}{} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{San Joaquin, CA} & \multicolumn{1}{l|}{077} & \multicolumn{1}{l|}{1} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Stanislaus, CA} & \multicolumn{1}{l|}{099} & \multicolumn{1}{l|}{} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Tulare, CA} & \multicolumn{1}{l|}{107} & \multicolumn{1}{l|}{} \\* \cmidrule(l){2-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{\multirow{4}{*}{\begin{tabular}[c]{@{}l@{}}Los Angeles-South \\ Coast Air Basin, CA\end{tabular}}} & \multicolumn{1}{l|}{Los Angeles, CA (p)} & \multicolumn{1}{l|}{037} & \multicolumn{1}{l|}{1} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Orange, CA} & \multicolumn{1}{l|}{059} & \multicolumn{1}{l|}{1} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Riverside, CA (p)} & \multicolumn{1}{l|}{065} & \multicolumn{1}{l|}{} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{San Bernardino, CA (p)} & \multicolumn{1}{l|}{071} & \multicolumn{1}{l|}{} \\* \cmidrule(l){2-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{Plumas County, CA} & \multicolumn{1}{l|}{Plumas, CA (p)} & \multicolumn{1}{l|}{063} & \multicolumn{1}{l|}{} \\* \midrule
\multicolumn{1}{|l|}{ID} & \multicolumn{1}{l|}{West Silver Valley, ID} & \multicolumn{1}{l|}{Shoshone, ID (p)} & \multicolumn{1}{l|}{079} & \multicolumn{1}{l|}{} \\* \midrule
\multicolumn{1}{|l|}{\multirow{2}{*}{OH}} & \multicolumn{1}{l|}{\multirow{2}{*}{Cleveland, OH}} & \multicolumn{1}{l|}{Cuyahoga, OH} & \multicolumn{1}{l|}{035} & \multicolumn{1}{l|}{} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Lorain, OH} & \multicolumn{1}{l|}{093} & \multicolumn{1}{l|}{} \\* \midrule
\multicolumn{1}{|l|}{\multirow{3}{*}{PA}} & \multicolumn{1}{l|}{Delaware County, PA} & \multicolumn{1}{l|}{Delaware, PA} & \multicolumn{1}{l|}{035} & \multicolumn{1}{l|}{} \\* \cmidrule(l){2-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{Lebanon County, PA} & \multicolumn{1}{l|}{Lebanon, PA} & \multicolumn{1}{l|}{075} & \multicolumn{1}{l|}{} \\* \cmidrule(l){2-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{Allegheny, PA} & \multicolumn{1}{l|}{Allegheny, PA} & \multicolumn{1}{l|}{005} & \multicolumn{1}{l|}{} \\* \bottomrule
\end{longtable}